---
title: 'Prediction: Titanic Survival'
author: "FrannieK"
date: "6 September 2016"
output: pdf_document
---

#1. Define the question #
Predict survival on the Titanic

#2. Define the ideal data set# 
Per given in Kaggle train set but with no missing values

#3. Determine what data I can access #
#4. Obtain the data #


```{r package}
library(xlsx)
```

```{r raw data}

dir = "C:/Users/Huran/Desktop/Data_Project/Proj_Titanic"
setwd(dir)
set.seed(2016)

tr = read.csv("Data/train.csv",header=TRUE, sep=",", na.strings =c("NA", "", " ")) # na.strings instructs what to interpre as missing values
test = read.csv("Data/test.csv", header=TRUE, sep =",", na.strings = c("NA", "", " "))

```
 
#5. Clean the data # 
First examine the raw data, get raw profile and find missing values.
* Upon examining, found _Survived_, _Pclass_ are categorical hence need to change into factors.
* Also performed the same to test data frame and added blank column for prediction later on.
* _Age_, _Cabin_ and _Embarked_ have missing values which will be filled with median value, NA and mode value respectively.



```{r clean data}
str(tr) # gets structure
str(test)
 
tr$Survived = as.factor(tr$Survived) # convert class of object
tr$Pclass = as.factor(tr$Pclass)

test[,"Survived"] = NA # adds a blank column
test$Pclass = as.factor(test$Pclass)

summary(tr) # descriptive profile
summary(test)

colSums(apply(tr,2, function(x) is.na(x))) # locate missing values
tr$Age[is.na(tr$Age)] = median(tr$Age[!is.na(tr$Age)]) # interchangeable with tr[["Age"]][n] or tr[,"Age"][n]
tr$Embarked[is.na(tr$Embarked)] = "S"

tr$Sex_ind = sapply(tr$Sex, function(x) if(x=="male") {1} else {2})
tr$Emb_ind = sapply(tr$Embarked, function(x) if(x=="C") {1} else if (x=="Q") {2} else {3})

test$Age[is.na(test$Age)] = median(test$Age[!is.na(test$Age)]) #do the same for test set 
test$Embarked[is.na(test$Embarked)]="S"

test$Sex_ind = sapply(test$Sex, function(x) if(x=="male") {1} else {2})
test$Emb_ind = sapply(test$Embarked, function(x) if(x=="C") {1} else if (x=="Q") {2} else {3})


write.csv(tr, file = "clean_data.csv", row.names=TRUE) # save clean train data in csv format
write.xlsx(test, file="clean_test.xlsx", row.names=TRUE)

```

#6. Exploratory data analysis
I'll take the clean data to SPSS for analytical steps. These include:
* mean, median, mode
* variance
* scatter plot
* hierachical clustering
* k means clustering
* principal component analysis
* ANOVA 

Features are as follows:
# mean
mean age is 30, 


# variance

...


